---
title: "Part 1 - data"
author: "jean wills"
date: "06/07/2020"
output: word_document
---

## Project: Portuguese Bank Marketing Data

# Step 1: read the data and look at the data structure

```{r}
# setwd("~/Users/jeanwills/Desktop/CKME136/")
# BM <- read.csv("/Users/jeanwills/Desktop/CKME136/1_data/bank_full.csv", header=T, sep = ";", stringsAsFactors = T, na.strings = "NA")
# look at the data structure 
# since we are using the smaller file, we will do the statistics on the smaller file
BM_mini <- read.csv("/Users/jeanwills/Desktop/CKME136/1_Data/bank.csv", header=T, sep = ";", stringsAsFactors = T, na.strings = "NA")
# let's check number of complete cases for no data missing at all -> no missing data!
BM<- BM_mini
sum(complete.cases(BM))
```

# step 2 - Look at the bank data summary

```{r}
summary(BM)
# we see that 7 attributes are numeric and the rest are now factors
# we will come back to this and change the classes where required
```

# Step 3a: plot the boxplots of the numeric data

```{r}
# age has outliers above ~70
par(mfrow=c(1,4))
boxplot(BM$age, main = "Client's age")
# balance has a large number of outliers above Q3
boxplot(BM$balance, main = "Balance")
# day has a large number of outliers above Q3
boxplot(BM$day, main = "day called on")
# duration has a large number of outliers above Q3
boxplot(BM$duration, main = "call duration (sec)")
```

# Step 3b

```{r}
par(mfrow=c(1,3))
# camapign has a large number of outliers above the Q3
boxplot(BM$campaign, main = "Boxplot of campaign")
# pdays has a VERY large number of outliers above Q3
boxplot(BM$pdays, main = "Boxplot of pdays")
# duration has a large number of outliers above Q3
boxplot(BM$previous, main = "Boxplot of previous")
```

# Step 4a: plot histograms to reveal skewness / normality

```{r}
par(mfrow=c(1,4))
# age looks skewed right
hist(BM$age, main = "Hist - Age", breaks = 5)
# balance is skewed right
hist(BM$balance, main = "Hist - Balance", breaks = 5)
# day somewhat skewed right
hist(BM$day, main = "Hist - day", breaks = 10)
# duration is skewed right
hist(BM$duration, main = "Hist - call duration", breaks = 5)
```

# Step 4b: plot histograms to reveal skewness / normality

```{r}
par(mfrow=c(1,3))
# campaign is skewed right - most data is in 1 day
hist(BM$campaign, main = "Histogram of campaign", breaks = 5)
# pdays skewed right
hist(BM$pdays, main = "Histogram of pdays", breaks = 5)
# previous skewed right
hist(BM$previous, main = "Histogram of previous", breaks = 5)
```

# 4c logs of the numeric data and replot the histograms...

probably won't use this as some data is negative and <1 and doing log(data) does NOT work for them but keep for now

```{r}
# NaNs produced.....need # >1.0 - balance and duration would need adjustments
BM_log<-BM
BM_log$age<-log(BM$age)
# balance has negative values 
BM_log$balance<-log(BM$balance)
BM_log$day<-log(BM$day)
# duration has zeros
BM_log$duration<-log(BM$duration)
#
par(mfrow=c(1,4))
hist(BM_log$age, main = "log(Age)", breaks = 5)
hist(BM_log$balance, main = "log(Balance)", breaks = 5)
hist(BM_log$day, main = "log(Day)", breaks = 5)
hist(BM_log$duration, main = "log(call duration)", breaks = 10)
```

# 4d logs of the numeric data and replot the histograms...

```{r}
# NaNs produced.....need # >1.0 - pdays and previous would need adjustments
BM_log$campaign<-log(BM$campaign)
# pdays has -1
BM_log$pdays<-log(BM$pdays)
# previous has 0 and 1
BM_log$previous<-log(BM$previous)
# str(BM_log)
par(mfrow=c(1,3))
hist(BM_log$campaign, main = "log(campaign)", breaks = 5)
hist(BM_log$pdays, main = "log(pdays)", breaks = 5)
hist(BM_log$previous, main = "log(previous)", breaks = 5)
# most seem normal except campaign is still right skewed
```

The logs work for most of the data but not all.....so leaving data as is

# Step 5a: Q-Q plots

```{r}
par(mfrow=c(1,2))
qqnorm(BM$age, main = "Norm Q-Q plot of Age")
qqline(BM$age)
qqnorm(BM$balance, main = "Norm Q-Q plot of balance")
qqline(BM$balance)
# both not normal
```

# Step 5b: Q-Q plots

```{r}
par(mfrow=c(1,2))
qqnorm(BM$day, main = "Norm Q-Q plot of day")
qqline(BM$day)
qqnorm(BM$duration, main = "Norm Q-Q plot of duration")
qqline(BM$duration)
# both not normal
```

# Step 5c: Q-Q plots

```{r}
par(mfrow=c(1,3))
qqnorm(BM$campaign, main = "Norm Q-Q plot of campaign")
qqline(BM$campaign)
qqnorm(BM$pdays, main = "Norm Q-Q plot of pdays")
qqline(BM$pdays)
qqnorm(BM$previous, main = "Norm Q-Q plot of previous")
qqline(BM$previous)
# all 3 not normal
```

# Step 6: Shapiro Tests for Normality on numeric data

```{r}
# IF p<0.05 then the numeric data is not normal and significant. 
# Shapiro requires dataset size under 5,000.
# All numeric attributes are NOT normal
shapiro.test(BM_mini$age)
shapiro.test(BM_mini$balance)
shapiro.test(BM_mini$day)
shapiro.test(BM_mini$duration)
shapiro.test(BM_mini$campaign)
shapiro.test(BM_mini$pdays)
shapiro.test(BM_mini$previous)
```

# Step 7a: test for correlations within the numeric attributes

```{r}
# Since we know the numeric data is not-normal, we use Spearman instead of Pearson method
# The correlation heat map is created
# Pearson method is default and p>0.05 means NOT correlated
# Spearman method - if p<0.05 means NOT correlated

# if p<0.05 then significant meaning correlated
# simple example with 2 variables
# if we do this, we need y as numeric 0/1
cor.test(BM$previous,BM$age, method="spearman")
cor.test(BM$previous,BM$age)
# also: cor(BM$previous,BM$age)
```

# Step 7b: Table of correlations for all data

```{r}
# Correlation test can also be considered a Feature Removal method
# test ALL data for correlations
library(lattice)
library(ggplot2)
BM_num<-BM
# num<- subset(BM_01, select = c("age", "balance", "day", "duration", "campaign", "pdays", "previous", "y"))
BM_num$job<- as.numeric(BM_num$job)  #12
BM_num$marital<- as.numeric(BM_num$marital) #4
BM_num$education<- as.numeric(BM_num$education) #4
BM_num$default<- as.numeric(BM_num$default) #2
BM_num$housing<- as.numeric(BM_num$housing)  #2
BM_num$loan<- as.numeric(BM_num$loan)  #2
BM_num$contact<- as.numeric(BM_num$contact) #3
BM_num$month<- as.numeric(BM_num$month)  #12
BM_num$poutcome<- as.numeric(BM_num$poutcome)  #4
# correlations data 
# Identify highly correlated features in caret r package
# ensure the results are repeatable
set.seed(12)
library(mlbench)
library(caret)
# leaving out y so only 16 not 17
corMatrix<-cor(BM_num[, c(1:16)])
print(corMatrix)
highCorr <- findCorrelation(corMatrix, cutoff=0.5)
print(highCorr)
```

# Step 7c: do a correlation heat map to visualize the data

## to do this, create all numeric data first (copied from Part 2):

```{r}
BM_n <- BM
BM_n$job<- as.numeric(BM_n$job)  #12
# marital: 1-single, 2-married, 3-divorced
BM_n$marital<- ifelse(BM_n$marital == c("single"), 1, 
                        ifelse(BM_n$marital== c("married"), 2, 3))
# education: 0:unknown, 1: primary, 2:secondary, 3:divorced
BM_n$education<- ifelse(BM_n$education == c("unknown"), 0, 
                          ifelse(BM_n$education == c("primary"), 1, 
                                 ifelse(BM_n$education == c("secondary"), 2, 3)))
# default, housing, loan: if yes then 0 else 1
# BM_num$housing<- as.numeric(BM_num$housing)  #2
BM_n$default<- ifelse(BM_n$default == c("yes"), 0, 1) #2
BM_n$housing<- ifelse(BM_n$housing == c("yes"), 0, 1) #2
BM_n$loan<- ifelse(BM_n$loan == c("yes"), 0, 1) #2
BM_n$contact<- as.numeric(BM_n$contact) #3
# month: jan:1, feb:2.....dec:12
BM_n$month<- ifelse(BM_n$month == "jan", 1, 
                     ifelse(BM_n$month == "feb", 2, 
                            ifelse(BM_n$month == "mar", 3,
                                   ifelse(BM_n$month == "apr", 4, 
                                          ifelse(BM_n$month == "may", 5, 
                                                 ifelse(BM_n$month == "jun", 6,
                                                        ifelse(BM_n$month == "jul", 7,
                      ifelse(BM_n$month == "aug", 8,
                            ifelse(BM_n$month == "sep", 9,
                                   ifelse(BM_n$month == "oct", 10,
                                          ifelse(BM_n$month == "nov", 11, 12)))))))))))
# poutcome: 0:unknown,other, 1:failure, 2: success
BM_n$poutcome<- ifelse(BM_n$poutcome == c("failure"), 1, ifelse(BM_n$poutcome== c("success"), 2, 0))  
# result: BM_num with only numeric data (NOT scaled)
# extra step for BM_mini
BM_num<-BM_n
rm(BM_n)
```

# Step 7c: do a correlation heat map to visualize the data

```{r}
library(plyr)
library(GGally)
library(ggplot2)
library(reshape2)
library(caret)
# BM_num was created above
# leaving out y so only 16 not 17
# this one prints extra info -  ggpairs(BM_num[, c(1,16)])
# require(scales)
bnk_core<- cor(BM_num[, c(1:16)])
bnk_melt<- melt(bnk_core, varnames=c("x", "y"),value.name="Correlation")
# summarize the correlation matrix
highlyCorrelated <- findCorrelation(bnk_core, cutoff=0.5)
# print indexes of highly correlated attributes
ggplot(bnk_melt, aes(x=x, y=y)) +
  geom_tile(aes(fill=Correlation)) + 
  scale_fill_gradient2(low="blue", mid="white", high="red", guide=guide_colorbar(ticks=FALSE, barheight=10),limits=c(-1,1)) +
  theme_minimal() +
  labs(x=NULL, y=NULL)
```

# Step 7d: scatterplot matrix of the numeric data - (this takes a bit of time)

```{r}
# BM_num created above 
library(psych)
# pairs works too
# top right part shows correlations, diagonal shows histograms, and bottom left shows the
# scatterplots, with the circles showing strength of correlation (circle~little, oval~lot)
# took out y in the end
pairs.panels(BM[c("age","balance","day","duration","campaign","pdays","previous")])
```

# Step 8a: Pearson chi-sq test for correlations of non-numeric data

This is a sample of what could be done if we did not do the heat map above

```{r}
#Also:
#chisq.test(table(BM$job, BM$marital))$expected
#and $expected shows what the results should look like if true under null hypothesis

# test for one attribute at a time against all the others for correlation
# if p<0.05 then not significant 
chisq.test(BM$job, BM$marital)
chisq.test(BM$job, BM$education)
# all p-values ~ zero - none significant
```

# Step 8b: Pearson chi-sq test for correlations of categorical data to the target variable

```{r}
chisq.test(BM$y, BM$job)
chisq.test(BM$y, BM$marital)
chisq.test(BM$y, BM$education)
chisq.test(BM$y, BM$housing)
chisq.test(BM$y, BM$loan)
chisq.test(BM$y, BM$contact)
chisq.test(BM$y, BM$month)
chisq.test(BM$y, BM$poutcome)
# all p-values ~ zero - this is not good! we want correlation!!
# this means that it may be difficult to get meaningful answers!
```

